---
title: "Time Series Assignment"
author: "Thanh-Tung LUONG"
date: "2024-04-15"
output: html_document
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,warning = FALSE, message = FALSE)
```
```{r}
rm(list=ls())
```

```{r}
library(dplyr)
library(psych)
library(ggplot2)
library(urca)
library(zoo)
library(exuber)
library(FinTS)
library(conflicted)
library(rugarch)
library(MSGARCH)
library(forecast)
library(rmgarch)
library(BEKKs)
library(moments)
```

--------------------------------------------------------------------------------

# ANALYSIS OF BTC's TIME SERIES

## Create variables and general statistical description

```{r}
btc <- read.csv("bitcoin-daily-2017.csv", header = TRUE, stringsAsFactors = TRUE)
btc$date <- as.Date(btc$date, format = "%Y-%m-%d")

#btc$Price <- na.approx(btc$Price)

btc <- btc %>%
  mutate(logPrice = log(Price))

btc <- btc %>%
  mutate(logReturn = c(NA,diff(logPrice))*100)
btc$logReturn[1] <- 1

btc$demean_ret = (btc$logReturn-mean(btc$logReturn))^2
print(btc)
```

```{r}
btc_stats <- describe(btc)
print(btc_stats)
```


## BTC's log price analysis

### Plot of the series and its ACF and PACF

```{r}
ggplot(data = btc, aes(x = seq_along(logPrice), y = logPrice)) +
  geom_line() + 
  labs(x="Day", y="Log Price", title="Log Bitcoin Price Plot")
```
The Time Series of log price of Bitcoin seems like a trend stationary process. We now carry out some tests to see the features of this time series.

```{r}
acf(x = btc$logPrice,lag.max =20,main="Autocorrelogram of log price of Bitcoin")
pacf(x = btc$logPrice,lag.max = 20,main="Partial Autocorrelogram of log price of Bitcoin")
```
As we can observe, the ACF is persistent over time (close to 1 and slightly decreasing), while the PACF is only significant at lag 1 and become insignificant from the second lag. We can conclude that this time series has characteristics of a difference stationary process.

### Unit Root Tests

#### ADF Test with deterministic trend and an intercept

```{r}
tt_drift_adf<-ur.df(y = btc$logPrice,type=c("trend"), lag=20, selectlags = c('AIC'))
summary(tt_drift_adf)
```

#### ADF Test with intercept, no trend

```{r}
drift_adf<-ur.df(y = btc$logPrice,type=c("drift"), lag=20, selectlags = c('AIC'))
summary(drift_adf)
```

#### ADF Test no intercept, no trend

```{r}
none_adf<-ur.df(y = btc$logPrice,type=c("none"), lag=20, selectlags = c('AIC'))
summary(none_adf)
```
From three tests above, the t-stats are all higher than 10% DF critical values. Thus, we cannot reject the null hypothesis (H0: not stationary). We come to the conclusion that log price of Bitcoin is not a second-order stationary process.

#### KPSS stationarity test around a deterministic trend

```{r}
kpss_tau<-ur.kpss(y=btc$logPrice,type="tau",lags="short")
summary(kpss_tau)
```

#### KPSS stationarity test around an intercept

```{r}
kpss_tau<-ur.kpss(y=btc$logPrice,type="mu",lags="short")
summary(kpss_tau)
```
The KPSS tests suggest that we reject the Null hypothesis (H0: the series is stationary) at 1 percent significance level, i.e: the series is not stationary.


#### Ljung-Box test for no autocorrelation

```{r}
LB=Box.test(btc$logPrice,lag=20,type="Ljung")
sprintf("Q-stat of the Ljung-box test: %.4f", LB$statistic)
```
Since the Q-stat is large than $\chi^{2}_{90%}(20)$ (8.55), we reject the null hypothesis of the Ljung-Box test (H0: no correlation until rank 20) at 1 percent significance level.


## BTC's log return analysis

### Plot of the series and its ACF and PACF 

```{r}
ggplot(data = btc, aes(x = seq_along(logReturn), y = logReturn)) +
  geom_line() + 
  labs(x="Day", y="Log Return", title="Bitcoin's Log Returns Plot")
```
```{r}
acf(x = btc$logReturn,lag.max=20,main="Bitcoin's log return autocorrelogram")
pacf(x = btc$logReturn,main="Bitcoin's log return partial autocorrelogram")
```
For autocorrelation, only the first lag is significant. For partial autocorrelation, we can see that it is only significant at the 10th lags.

Again, let's carry out the unit root test to see the feature of this series.

### Unit Root Tests

#### ADF Test with deterministic trend and an intercept

```{r}
tt_drift_adf<-ur.df(y = btc$logReturn,type=c("trend"), lag=20, selectlags = c('AIC'))
summary(tt_drift_adf)
```

#### ADF Test with intercept, no trend

```{r}
drift_adf<-ur.df(y = btc$logReturn,type=c("drift"), lag=20, selectlags = c('AIC'))
summary(drift_adf)
```

#### ADF Test no intercept, no trend

```{r}
none_adf<-ur.df(y = btc$logReturn,type=c("none"), lag=20, selectlags = c('AIC'))
summary(none_adf)
```

As observed through all conducted tests, the t-stats are all lower than critical values for 1% significance level. Thus, we reject the null hypothesis of the ADF unit root test, i.e: there is stationarity in the series.

#### KPSS stationarity test around a deterministic trend

```{r}
kpss_tau<-ur.kpss(y=btc$logReturn,type="tau",lags="short")
summary(kpss_tau)
```

#### KPSS stationarity test around an intercept

```{r}
kpss_tau<-ur.kpss(y=btc$logReturn,type="mu",lags="short")
summary(kpss_tau)
```
The KPSS tests suggest that we cannot reject the Null hypothesis (H0: the series is stationary), since the p-values are all smaller than 10 percent significance level.


#### Ljung-Box test for no autocorrelation

```{r}
LB=Box.test(btc$logReturn,lag=20,type="Ljung")
sprintf("Q-stat of the Ljung-box test: %.4f", LB$statistic)
```
Since the Q-stat is large than $\chi^{2}_{90%}(20)$ (8.55), we reject the null hypothesis of the Ljung-Box test (H0: no correlation until rank 20) at 1 percent significance level.


## Test For Bubbles

Since there is no dividend for Bitcoin, we are going to test the speculative bubble for the log price series of Bitcoin 

```{r}
adf2<-ur.df(y = btc$logPrice,type=c("drift"), lag=12, selectlags = c('AIC'))
summary(adf2)
```

The number of lag here should be 2, according to AIC.

```{r}
psy_minw(btc$logPrice) # recommended window for the initial forward recursive estimation 
sprintf("The recommended window for the initial forward recursive estimation is  %f",psy_minw(btc$logPrice))
radf_btc <-radf(btc$logPrice,lag=2) # radf uses psy_minw(x) by default
sprintf("The SADF test statistics is equal to %f",radf_btc$sadf)
```
```{r}
radf_crit$n119$sadf_cv
```
The SupADF statistics $SADF = 1.93$ is higher than the three critical values. We reject the null hypothesis of a unit root in favor of an explosive behavior at 1% significance level, i.e there is an evidence of a bubble. We can check again with simulated critical values. 

```{r}
x<-btc$logPrice
mc_critical_values <- radf_mc_cv(n=length(x), nrep=1000,seed = 123)
summary(radf_btc,mc_critical_values)
```

We use this simulated critical value for the unit root test:
```{r}
diagnostics(radf_btc,mc_critical_values,option = "sadf")
```

And for the bubble detection:
```{r}
bubble_date<-datestamp(radf_btc,cv=mc_critical_values,min_duration = psy_ds(x),option = "sadf")
bubble_date
```

```{r}
autoplot_x_sadf<-autoplot(radf_btc,option="sadf",cv=mc_critical_values,min_duration = psy_ds(x))+labs(title="Dates of the speculative bubble - sadf test")
autoplot_x_sadf
```

## Modeling of the returns and the conditional variances

### GARCH Models

```{r}
plot(btc$date, btc$logReturn,type='l',col='blue',xlab = "",ylab = "",xaxt="n",main = "BTC's log returns")
axis.Date(1,at=seq(gold_df$date[1],tail(btc$date,1),by="3 months"),format = "%d/%m/%y",las=2)
```
```{r}
plot(btc$date, btc$demean_ret,type='l',col='blue',xlab = "",ylab = "",xaxt="n",main = "BTC's demeaned returns")
axis.Date(1,at=seq(btc$date[1],tail(btc$date,1),by="3 months"),format = "%d/%m/%y",las=2)
```
```{R}
acf(x = btc$demean_ret,lag.max =20,main="Autocorrelogram BTC squared demeaned log return")
```
```{r}
ArchTest(btc$logReturn,lags = 2,demean=TRUE)
```

pvalue is close to 0 => we reject the null of no ARCH effect

#### ARIMA(1,0,1)+GARCH(1,1) without variance targeting and a Gaussian distribution

```{r}
args(ugarchspec)
spec = ugarchspec()
fit_btcret=ugarchfit(spec = spec,data = btc$logReturn,solver = "hybrid")
show(fit_btcret)
plot(fit_btcret, which="all")
```
```{r}
plot(btc$date,fit_btcret@fit$var,type="l",col="blue",main="Conditional Variance of BTC's log return",xlab="",ylab="")

# conditional standard deviation estimated from the model
plot(btc$date,fit_btcret@fit$sigma,type="l",col="blue",main="Conditional S.D of Gold's log return",xlab="",ylab="")
```

#### AR(1) + GARCH(1,1) + Gaussian distribution

```{r}
spec_AR1_garch11= ugarchspec(mean.model = list(armaOrder=c(1,0)))
show(spec_AR1_garch11)
fit_AR1_garch11=ugarchfit(spec = spec_AR1_garch11,data = btc$logReturn,solver = "hybrid")
show(fit_AR1_garch11)
plot(btc$date,fit_AR1_garch11@fit$sigma,type="l",col="blue",main="Conditional S.D of BTC's log return",xlab="",ylab="")
```

#### Exponential GARCH + Normal

```{r}
spec_egarch_Mnorm = ugarchspec(mean.model=list(armaOrder=c(1,0),archm=TRUE,archpow=1),variance.model = list(model="sGARCH"),distribution.model = "norm")
show(spec_egarch_Mnorm)
fit_egarch_Mnorm = ugarchfit(spec = spec_egarch_Mnorm,data = btc$logReturn,solver = "hybrid")
show(fit_egarch_Mnorm)
```

#### GJR model+ AR1 + Gaussian distribution
```{r}
spec_AR1_GJR_norm = ugarchspec(variance.model = list(model="gjrGARCH"),distribution.model = "norm",mean.model = list(armaOrder=c(1,0)))
show(spec_AR1_GJR_norm)

fit_AR1_GJR_norm=ugarchfit(spec = spec_AR1_GJR_norm,data = btc$logReturn,solver = "hybrid")
show(fit_AR1_GJR_norm)
```

```{r}
plot(btc$date,fit_AR1_GJR_norm@fit$sigma,type="l",col="blue",main="Conditional S.D of BTC's log return - GJR",xlab="",ylab="")
matplot(x = btc$date, cbind(fit_AR1_garch11@fit$sigma,fit_AR1_GJR_norm@fit$sigma),lty=1,
        type = "l",col=c('red',"blue"),xlab="time",ylab = "sigma",main="Conditional S.D of BTC's log return with and without an asymmetry effect")

legend("topleft",legend=c("garch11","gjr11"),col=c('red',"blue"),lty=1)
```

#### Exponential GARCH model + Gaussian distribution
```{r}
spec_egarch = ugarchspec(mean.model=list(armaOrder=c(1,0)),variance.model = list(model="eGARCH"))
show(spec_egarch)

fit_egarch = ugarchfit(spec = spec_egarch,data = btc$logReturn,solver = "hybrid")
show(fit_egarch)
plot(fit_egarch@fit$sigma,type="l",col="blue",main="Conditional Standard Deviation of BTC's log return - eGarch",xlab="",ylab="")

matplot(x = btc$date, cbind(fit_AR1_garch11@fit$sigma,fit_AR1_GJR_norm@fit$sigma,fit_egarch@fit$sigma),lty=1,
        type = "l",col=c('red',"blue","green"),xlab="time",ylab = "sigma",main="Conditional Standard deviation with and without an asymmetry effect")

legend("topleft",legend=c("garch11","gjr11","egarch11"),col=c('red',"blue","green"),lty=1)

```
gamma1 (coefficient of the asymmetric effect) is significant, suggesting evidence of an asymmetric effect of negative shocks on return.

#### GARCH(1,1) with a Student distribution for the standardized residuals

```{r}
spec_AR1_St = ugarchspec(mean.model = list(armaOrder=c(1,0)),distribution.model="std")
show(spec_AR1_St)

fit_AR1_St=ugarchfit(spec = spec_AR1_St,data = btc$logReturn,solver = "hybrid")
show(fit_AR1_St)

plot(btc$date,fit_AR1_St@fit$var,type="l",col="blue",main="Conditional Standard Deviation of BTC's log return",xlab="",ylab="")
```

The information criteria are lower than with a Gaussian distribution, an evidence in favor of the Student distribution

#### AR(1) + GARCH(1,1) + GED distribution

```{r}
spec_AR1_ged = ugarchspec(mean.model = list(armaOrder=c(1,0)),distribution.model="ged")
show(spec_AR1_ged)

fit_AR1_ged=ugarchfit(spec = spec_AR1_ged,data = btc$logReturn,solver = "hybrid")
show(fit_AR1_ged)

plot(btc$date,fit_AR1_ged@fit$var,type="l",col="blue",main="Conditional Standard Deviation of BTC's log return",xlab="",ylab="")
```

#### AR(1) + GARCH(1,1) + Skewed-Student distribution

```{r}
spec_AR1_sSt = ugarchspec(mean.model = list(armaOrder=c(1,0)),distribution.model="sstd")
show(spec_AR1_sSt)

fit_AR1_sSt=ugarchfit(spec = spec_AR1_sSt,data = btc$logReturn,solver = "hybrid")
show(fit_AR1_sSt)

plot(btc$date,fit_AR1_sSt@fit$var,type="l",col="blue",main="Conditional Standard Deviation of BTC's log return",xlab="",ylab="")
```

#### AR(1) + GARCH(1,1) + skewed GED distribution

```{r}
spec_AR1_sged = ugarchspec(mean.model = list(armaOrder=c(1,0)),distribution.model="sged")
show(spec_AR1_sged)

fit_AR1_sged=ugarchfit(spec = spec_AR1_sged,data = btc$logReturn,solver = "hybrid")
show(fit_AR1_sged)

plot(btc$date,fit_AR1_sged@fit$var,type="l",col="blue",main="Conditional Standard Deviation of BTC's log return",xlab="",ylab="")
```

The information criteria are lower than with a Gaussian distribution, evidence in favor of the skewed-GED distribution

#### GJR model+ AR1 + skew-student distribution
```{r}
spec_AR1_GJR_sstd = ugarchspec(variance.model = list(model="gjrGARCH"),distribution.model = "sstd",mean.model = list(armaOrder=c(1,0)))
show(spec_AR1_GJR_sstd)

fit_AR1_GJR_sstd=ugarchfit(spec = spec_AR1_GJR_sstd,data = btc$logReturn,solver = "hybrid")
show(fit_AR1_GJR_sstd)

plot(btc$date,fit_AR1_GJR_sstd@fit$sigma,type="l",col="blue",main="Conditional Standard Deviation of BTC's log return - GJR",xlab="",ylab="")
```

#### GJR model+ AR1 + skew-ged distribution
```{r}
spec_AR1_GJR_sged = ugarchspec(variance.model = list(model="gjrGARCH"),distribution.model = "sged",mean.model = list(armaOrder=c(1,0)))
show(spec_AR1_GJR_sged)

fit_AR1_GJR_sged=ugarchfit(spec = spec_AR1_GJR_sged,data = btc$logReturn,solver = "hybrid")
show(fit_AR1_GJR_sged)

plot(btc$date,fit_AR1_GJR_sged@fit$sigma,type="l",col="blue",main="Conditional Standard Deviation of BTC's log return - GJR",xlab="",ylab="")
```

#### Exponential GARCH model + Skew Student
```{r}
specegarchsstd = ugarchspec(mean.model=list(armaOrder=c(1,0)),variance.model = list(model="eGARCH"),distribution.model = "sstd")
show(specegarchsstd)

fitegarchsstd=ugarchfit(spec = specegarchsstd,data = btc$logReturn,solver = "hybrid")
show(fitegarchsstd)
plot(fitegarchsstd@fit$sigma,type="l",col="blue",main="Conditional Standard Deviation of BTC's log return GJR",xlab="",ylab="")
```

#### exponential GARCH model + Skew Student + ARCH-in-mean

GARCH in Mean introduces a trade-off between risk and return.

```{r}
spec_egarchM_sstd = ugarchspec(mean.model=list(armaOrder=c(1,0),archm=TRUE,archpow=1),variance.model = list(model="eGARCH"),distribution.model = "sstd")
show(spec_egarchM_sstd)
fit_egarchM_sstd=ugarchfit(spec = spec_egarchM_sstd,data = btc$logReturn,solver = "hybrid")
show(fit_egarchM_sstd)
```


### Markov switching GARCH

```{r}
MS_spec<-CreateSpec()
print(MS_spec)

#Estimation of an AR1 model
AR1<-Arima(y = btc$logReturn,order=c(1,0,0))

#Estimation of the MS-GARCH
MS_Garch_Fit<-FitML(MS_spec,data=AR1$residuals)

# We display the estimation results
summary(MS_Garch_Fit)

ExtractStateFit(MS_Garch_Fit)
```

```{r}
# Estimation of the unconditional volatity level for each state
sapply(ExtractStateFit(MS_Garch_Fit),UncVol)

#Display of transition matrix
TransMat(MS_Garch_Fit)

#Computation and plot of conditional probability
P<-State(object = MS_Garch_Fit)
```

#### Specification with asymmetric effect

```{r}
specgjr<-CreateSpec(variance.spec = list(model=c("gjrGARCH","gjrGARCH")),distribution.spec = list(distribution=c("norm","norm")))
print(specgjr)
MS_Garchgjr_Fit<-FitML(specgjr,data=AR1$residuals)
summary(MS_Garchgjr_Fit)
```
No significant asymmetric effect for state 2

```{r}
specgjr<-CreateSpec(variance.spec = list(model=c("gjrGARCH","sGARCH")),distribution.spec = list(distribution=c("norm","norm")))
print(specgjr)
MS_Garchgjr_Fit<-FitML(specgjr,data=AR1$residuals)
summary(MS_Garchgjr_Fit)
```


#### Specification with non gaussian distributions skewed student

```{r}
specgjr_ssd<-CreateSpec(variance.spec = list(model=c("gjrGARCH","sGARCH")),distribution.spec = list(distribution=c("sstd","sstd")),constraint.spec=list(regime.const=c("nu","xi")))
MS_Garchgjr_Fit_ssd<-FitML(specgjr_ssd,data=AR1$residuals)
summary(MS_Garchgjr_Fit_ssd)
```












